---
title: "DS2003: Interactive Maps"
output:
     html_document:
          number_sections: yes
          toc: TRUE
          toc_depth: 4
          toc_float: true
          toc_collapsed: true
          theme: journal
          code_folding: hide
runtime: shiny          
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval=FALSE)
library(sf)
library(tidyverse)
library(tigris) #download shp file from U.S. census website
library(tmaptools)
library(tmap)
```

# Interactive Maps

- Interactive maps allow us to do:
     - Zooming
     - Panning around
     - Hovering
     - Changing map types, etc.


## Example Data

- Again, I use a dataset (created on December 6th, 2020) on state voting counts from 
[NBC news](https://www.nbcnews.com/politics/2020-elections/president-results),
along with States Latitudes and Longitudes from 
[Google](https://developers.google.com/public-data/docs/canonical/states_csv).

- I create voting percentages within the range of [0.4, 0.6] so that I could eventually color the map red and blue in a manner that closely reflected population political leanings.


```{r, eval=TRUE}
US2020_Election <- readxl::read_excel("Presidential_Results_Dec6_2020.xlsx") %>%
  mutate(Total = Trump+Biden) %>%
  mutate(Trump_Percent = 
           ifelse(Trump/Total>0.6, 0.6, ifelse(Trump/Total<0.4, 0.4, Trump/Total)))

US2020_Election %>% head()
```


# Package: `tmap`

- `tmap` is a powerful and flexible map-making package. It has a concise syntax that allows for the creation of attractive maps with minimal code which will be familiar to `ggplot2` users. 

- For more info, see [tmap-getstarted](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html), [Making maps with R](https://geocompr.robinlovelace.net/adv-map.html#fig:tmshape), as well as [Tennekes (2018)](https://www.jstatsoft.org/article/view/v084i06).

- Let's use the example 
  
```{r, eval=FALSE}
library(tigris) #download shp file from U.S. census website
library(tmaptools)
library(tmap)
library(sf)

download.file("http://www2.census.gov/geo/tiger/GENZ2015/shp/cb_2015_us_state_20m.zip", destfile = "states.zip")
unzip("states.zip")
```

```{r, eval=TRUE}
usmap <- st_read("cb_2015_us_state_20m.shp", stringsAsFactors = FALSE)
```

- Merge data and draw a choropleth map with package `tmap`.

```{r, eval=TRUE}
usdata <- inner_join(usmap, US2020_Election, by=c("STUSPS" = "State_Abbreviation"))
usdata2 <- usdata %>% filter(!STUSPS %in% c("PR", "HI", "AK")) # delete Puerto Rico, Hawaii, and Alaska
tm_shape(usdata2) + tm_polygons("Trump_Percent", id="STUSPS", palette = "Reds", title = "Rep %")
```

- Let's create an interactive map. Use the tmap interactive view mode.

```{r, eval=FALSE}
tmap_mode("view") # tmap view mode
tmap_last() # hovering, clicking, zooming, changing map types, saving files

# save the result as html
class_map <- tmap_last()
tmap_save(class_map, "class_tmap.html") 
```

# Package: `plotly`

- Let's use [Plotly](https://plot.ly/r/) that allows you to easily convert your ggplot2 graphics to an interactive web-based version. 

## `ggplotly()`

- `ggplotly()` takes existing `ggplot2` objects and converts them into interactive `plotly` graphics. That is, `ggplotly()` converts your static plots to an interactive web-based version!

- This makes it easy to create interactive figures because we are already familiar to the `ggplot2` syntax. 

```{r, eval=TRUE}
library(plotly)

g <-  ggplot(usdata2) + geom_sf(aes(fill=Trump_Percent)) + theme_bw() +
  scale_fill_gradient(low="beige", high="red", limits=c(.4,.6), name='Rep %')
ggplotly(g)
```

## `plot_ly()`

- `plot_ly()` is the base `plotly` command to initialize a plot from a dataframe, similar to `ggplot()` from ggplot2. For more info, see [https://plotly.com/r/choropleth-maps/](https://plotly.com/r/choropleth-maps/).

```{r, eval=TRUE}
# Load the state map data 
states <- ggplot2::map_data("state") # from ggplot2

US2020_Election$region <- tolower(US2020_Election$State_Name) # small letters

# Merge datasets
mapdata <- merge(states, US2020_Election, by = "region")

# Add hover text
mapdata$hover <- with(mapdata, paste(State_Name, '<br>', "Trump Percent:", format(round(Trump_Percent, 2), 2)))

# Make state borders red
borders <- list(color = toRGB("red"))

# Change some mapping options
map_options <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

plot_ly(mapdata, z = ~Trump_Percent, text = ~hover, locations = ~State_Abbreviation, 
        type = 'choropleth', locationmode = 'USA-states', 
        color = ~Trump_Percent, colors = 'Reds', marker = list(line = borders)) %>%
  layout(geo = map_options) %>% 
  colorbar(title = "Rep %") # change the title of colorbar
```

# `plotly` with Shiny Apps

- Use `renderPlotly()` in the `server` argument, and use `plotlyOutput()` in the `ui` argument.

```{r, eval=TRUE}
library(plotly)
library(shiny)
library(shinythemes)

states <- map_data("state") # from ggplot2

US2020_Election$region <- tolower(US2020_Election$State_Name) # small letters

# Merge datasets
mapdata <- merge(states, US2020_Election, by = "region")

# Add hover text
mapdata$hover <- with(mapdata, paste(State_Name, '<br>', "Trump Percent:", format(round(Trump_Percent, 2), 2)))

# Make state borders red
borders <- list(color = toRGB("red"))

# Change some mapping options
map_options <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)



ui = fluidPage(theme = shinytheme("united"),
               sliderInput(inputId = "percent", "Rep %", min = 0.4, max = 0.6, step = 0.05, value = 0.5),
               plotlyOutput(outputId = "map")
)

server = function(input, output) {
  
  output$map = renderPlotly({
    
    plot_ly(mapdata[mapdata$Trump_Percent <= input$percent, ], z = ~Trump_Percent, 
            text = ~hover, locations = ~State_Abbreviation, 
            type = 'choropleth', locationmode = 'USA-states', 
            color = ~Trump_Percent, 
            colors = 'Reds', 
            marker = list(line = borders)) %>%
      layout(title = 'Trump 2020', geo = map_options) %>% 
      colorbar(title = "Rep %", limits = c(0.4, 0.6)) # fix the range of colorbar
  })
}

shinyApp(ui, server, options = list(height=600))       
```


