---
title: "Plotly Lab"
week: 9 
---

```{r}
library(tidyverse)
library(plotly)
```

```{r}
data <- as.data.frame(iris)
head(data)
```

# create a scatterplot of petal length vs petal width colored by species using ggplot2. Set you colors manually. 

```{r}

```

# Convert your plot above to an interactive graph using `ggplotly`

```{r}

```

# Now code the the above plot directly in `plot_ly`

1. Manipulate data as needed
```{r}

```

2. initialize your plotly plot
```{r}

```

3. add your Setosa trace 
```{r}

```

4. Add your versicolor trace
```{r}

```

5. Add your virginica trace
```{r}

```

6. Update the tooltip to read petal width: x.x petal length: x.x
```{r}

```

7. Add graph title and remove the gridlines
```{r}

```

8. Lighten the marker, create a dark outline of the same color
```{r}

```

Homework:

9. Take the above graph and make it 3D by adding Sepal Length on the z-axis. 
```{r}

```

10. Create grouped boxplots where each iris dataset measure in on the x-axis with 3 boxplots for each measure representing the species. Use your own cohesive color palette using coolors.co. Don't forget to include all the parts of a chart. 

```{r}

```

