---
title: 'DS2003: Interactive Plots in plotly'
author: "Natalie Kupperman"
week: 10
class: 19
---
## Let's work through some of the questions that came up last week...

```{r}
library(tidyverse)
library(plotly)
```

```{r}
# read in 'wine.csv' data
wine <- read.csv('wine.csv')

head(wine)
```

```{r}
wine$Type <- as.factor(wine$Type)
```

## Set bins of `plot_ly` histogram

* Use the argument `bingroup`

### Flavanoids by Type Histogram

```{r}
Type1 <- wine %>% filter(Type == 1)
Type2 <- wine %>% filter(Type == 2)
Type3 <- wine %>% filter(Type == 3)

```

```{r}

g7 <- plot_ly(alpha = 0.4) %>%
  add_histogram(x = ~Type1$Flavanoids,
                bingroup = 1, 
                name = 'Type 1',
                opacity = 0.6, 
                marker = list(color = "#00AFBB",
                              alpha = 0.6,
                              line = list(color = "lightgray",
                              width = 2))) %>%
  add_histogram(x = ~Type2$Flavanoids,
                bingroup = 1, 
                name = 'Type 2',
                opacity = 0.6,
                marker = list(color = "#E7B800",
                              line = list(color = "lightgray",
                              width = 2))) %>%
  add_histogram(x = ~Type3$Flavanoids, 
                bingroup = 1,
                name = 'Type 3',
                opacity = 0.6,
                marker = list(color = "#FC4E07",
                              line = list(color = "lightgray",
                              width = 2))) %>%
  layout(barmode = 'overlay',
         title = 'Histogram of flavanoids by soil type',
         xaxis = list(title = 'flavanoids',
                      zeroline = FALSE),
         yaxis = list(title = 'count'))
g7



g7_2 <- wine %>% plot_ly(alpha = 0.4, type = "histogram", marker = list(width = 5)) %>% 
  add_histogram(x = ~Flavanoids , color = ~Type, bingroup = 1,) %>%
  layout(barmode = 'overlay',
         xaxis = list(title = 'flavanoids',
                      zeroline = FALSE),
         yaxis = list(title = 'count'))
g7_2

```
## Scatter plot: 

```{r}
g8 <- plot_ly(type = 'scatter', mode = 'markers') 

g8 <- g8 %>%
  add_trace(
    x = Type1$Flavanoids,
    y = Type1$Proline,
    opacity = 0.5,
    marker = list(
      color = "#00AFBB",
      size = 7,
      line = list(
        color = 'lightgray',
        width = .5
      )
    ),
    name = 'Type 1'
  ) 
g8 <- g8 %>%
  add_trace( x = Type2$Flavanoids, y = Type2$Proline,
    marker = list(color = "#E7B800", size = 7,
      line = list(color = 'lightgray', width = .5)
    ),
    name = 'Type 2'
  )

g8 <- g8 %>%
  add_trace(x = Type3$Flavanoids, y = Type3$Proline,
    marker = list(color = "#FC4E07", size = 7,
      line = list(color = 'lightgray', width = .5)),
    name = 'Type 3'
  )

g8 <- g8 %>%
  layout(title = 'Scatterplot of Flavanoids & Proline',
         xaxis = list(title = 'flavanoids',
                      zeroline = FALSE),
         yaxis = list(title = 'proline'))
g8


g8_2 <- wine %>% plot_ly(type = 'scatter', mode = 'markers') %>% 
  add_trace(x = ~Flavanoids, y = ~Proline, color = ~Type) %>% 
layout(title = 'Scatterplot of Flavanoids & Proline',
         xaxis = list(title = 'flavanoids',
                      zeroline = FALSE),
         yaxis = list(title = 'proline'))
g8_2


```

### setting transparency

```{r}
fig2 <- plot_ly(type = 'scatter', mode = 'markers') 

fig2 <- fig2 %>%
  add_trace(
    x = jitter(Type1$Flavanoids),
    y = jitter(Type1$Proline),
    opacity = 0.5,
    marker = list(
      color = 'rgba(0,0,0,1)',
      size = 7,
      line = list(
        color = 'lightgray',
        width = .5
      )
    ),
    name = 'Type 1'
  ) 

fig2 <- fig2 %>%
  add_trace(
    x = jitter(Type2$Flavanoids),
    y = jitter(Type2$Proline),
    marker = list(
      color = 'rgba(0,0,0,.5)',
      size = 7,
      line = list(
        color = 'lightgray',
        width = .5
      )
    ),
    name = 'Type 2'
  )

fig2 <- fig2 %>%
  add_trace(
    x = jitter(Type3$Flavanoids),
    y = jitter(Type3$Proline),
    marker = list(
      color = 'rgba(0,0,0,.1)',
      size = 7,
      line = list(
        color = 'lightgray',
        width = .5
      )
    ),
    name = 'Type 3'
  )

fig2 <- fig2 %>%
  layout(title = 'Scatterplot of Flavanoids & Proline',
         xaxis = list(title = 'flavanoids',
                      zeroline = FALSE),
         yaxis = list(title = 'proline'))
fig2



```


## Custom controls in Plotly (hint: may be useful on final project)

For each feature, we will take a few minutes to look through the documentation. We will then go over some of the code together.

1. [Custom Buttons](https://plotly.com/r/custom-buttons/)

2. [Dropdown Events](https://plotly.com/r/dropdowns/)

3. [Sliders](https://plotly.com/r/range-slider/)

4. [3D Scatter - Life Expectancy](https://plotly.com/r/3d-scatter-plots/)

## Examples of Animated Plots

What's the difference between interactive and animated plots?  

1. [Racing Bar Chart](https://observablehq.com/@d3/bar-chart-race)  

2. [Animated Scatter](https://observablehq.com/@mbostock/the-wealth-health-of-nations)  

3. [Animated Line Graph](https://observablehq.com/@d3/connected-scatterplot)  

4. [Zoomable Circle](https://observablehq.com/@d3/zoomable-circle-packing)


