---
title: "Static Maps - Packages"
class: "DS2002: Communicating with Data"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval=FALSE)
library(tidyverse)
library(sf)
library(tigris) #download shp file from U.S. census website
library(tmaptools)
library(tmap)
library(ggmap)
```

# Mapping packages

## Package: `ggmap`

## Obtain Google Maps API Keys


```{r, eval=FALSE}
library(ggmap)

# Confirm that you have a registered google key
if (!has_google_key()) {stop("You need a google API key for this exercise")}

# If it error please follow the instructions with maps intro

```


## Plot Maps 

- Use package `ggmap`. 

- Three functions:
 
   - `qmap()` plots a quick map

   - `get_map()` gets map data for a location.

   - `ggmap()` plots stored map data

```{r}
register_google()

# Use qmap (quick map) to draw a city map
qmap("Washington DC", zoom=10)
```

```{r}
# Try different values for zooming
qmap("Washington DC", zoom=15)
```

```{r}
qmap("Washington DC", zoom=18)
```

```{r}
qmap("Washington DC", zoom=7)
```

```{r}
# Get the same map with get_map & save it to a variable
dc_map <- get_map("Washington DC", zoom=10)
dc_map
```

```{r}
# And plot it with ggmap
ggmap(dc_map) + xlab("Longitude") + ylab("Latitude")
```

## Geocode Points

- Pinpoint specific locations on the map.

- Geocoding converts a character string to a longitude and latitude location.

- `geocode()` takes a location name as input and returns the latitude and longitude. 
    
```{r, eval=TRUE}
# Geocode locations of interest
dc <- geocode("Washington DC")
dc # the longitude and latitude values
```

```{r}
wh <- geocode("White House")
wh
```

```{r}
uva <- geocode("University of Virginia")
uva
```

```{r}
# Map a point
uva_map <- ggmap(get_map(uva))
uva_map
```

## Change Map Types

- Terrain Map is the default map, and includes major place names and roads, but it also includes geographic terrain features. 

- Use `?get_map` to see available types. 

```
maptype = c("terrain", "terrain-background", "satellite", "roadmap", "hybrid",
    "toner", "watercolor", "terrain-labels", "terrain-lines", "toner-2010", "toner-2011",
    "toner-background", "toner-hybrid", "toner-labels", "toner-lines", "toner-lite")
```  

- Different types of maps are useful in different situations. For example,

    - Terrain maps provide useful context. But it has too much info to draw data points on the map.
    - The toner map is useful if we need to put a lot of data on the top of the map.
    
```{r}
# Try different map types
ggmap(get_map(uva, maptype = "terrain", zoom=13))
```

```{r}
ggmap(get_map(uva, maptype = "terrain-lines", zoom=13)) 
```

```{r}
ggmap(get_map(uva, maptype = "satellite", zoom=13))
```

```{r}
ggmap(get_map(uva, maptype = "roadmap", zoom=13))
```

```{r}
ggmap(get_map(uva, maptype = "hybrid", zoom=13))
```

```{r}
ggmap(get_map(uva, maptype = "toner-lite", zoom=13))
```

## Add Points on Map

```{r, eval=TRUE}
# Geocode locations of interest
eus <- geocode("West Virgnia")
dc <- geocode("Washington DC")

# Create a map of the Eastern US
ggmap(get_map(eus))
eus_map <- ggmap(get_map(eus, zoom=5))
eus_map
```


```{r, eval=TRUE}
# Add a location on the map
ggmap(get_map(eus, zoom=5)) + geom_point(data=dc, mapping=aes(x=lon, y=lat), color="red")
```

```{r, eval=TRUE}
# Add multiple datapoints
loc_names <- c("White House", "University of Virginia", "Statue of Liberty", "Boston University")
loc_values <- geocode(loc_names)
loc_dat <- data.frame(name=loc_names, lat=loc_values$lat, lon=loc_values$lon)

# Add multiple locations on the map
eus_map + geom_point(data=loc_dat, mapping=aes(x=lon, y=lat), color="red")
```

```{r, eval=TRUE}
# Add their names as text and adjust the label position
eus_map + geom_point(data=loc_dat, mapping=aes(x=lon, y=lat), color="red") +
  geom_text(data=loc_dat, mapping=aes(x=lon, y=lat, label=name), nudge_y=-0.5, nudge_x=1, size=2, color="red")
```

```{r, eval=TRUE}
# Try a different map type.
ggmap(get_map(eus, zoom=5, maptype = "terrain-background")) + 
  geom_point(data=loc_dat, mapping=aes(x=lon, y=lat), color="red") +
  geom_text(data=loc_dat, mapping=aes(x=lon, y=lat, label=name), nudge_y=-0.5, nudge_x=1, size=2, color="red")
```

```{r}
ggmap(get_map(eus, zoom=5, maptype = "toner-background")) +
  geom_point(data=loc_dat, mapping=aes(x=lon, y=lat), color="red") +
  geom_text(data=loc_dat, mapping=aes(x=lon, y=lat, label=name), nudge_y=-0.5, nudge_x=1, size=2, color="red")
```

```{r}
ggmap(get_map(eus, zoom=5, maptype = "watercolor")) + 
  geom_point(data=loc_dat, mapping=aes(x=lon, y=lat), color="red") +
  geom_text(data=loc_dat, mapping=aes(x=lon, y=lat, label=name), nudge_y=-0.5, nudge_x=1, size=2, color="red")
```


# Package: `ggplot2`

## Get state data

- `ggplot2` provides basic maps that allow for advanced visualizations.

- The `map_data()` function from `ggplot2` gets the set of points that define a region's boundaries.

- If you don't define the `group` aesthetic, you're drawing one single polygon that connects all the data points rather than a set of 48 different polygons from `state` map data.

```{r, eval=TRUE}
# Load the state map data 
states <- map_data("state")

# Inspect the state data
knitr::kable(head(states)) # The data points define each state as an individual polygon.

# Plot the state data
ggplot(data=states, mapping=aes(x=long, y=lat)) + 
  geom_polygon()
```

```{r, eval=TRUE}
# Improve the plot
ggplot(data=states, mapping=aes(x=long, y=lat, group=group, fill = group)) + 
  geom_polygon() +
  coord_map() + # fix the coordinate system
  theme_bw() + # change the background
  labs(x='Longitude', y = 'Latitude') # add x and y labels
```

## Create Choropleth Maps

- I used a dataset (created on December 6th, 2020) on state voting counts from 
[NBC news](https://www.nbcnews.com/politics/2020-elections/president-results),
along with States Latitudes and Longitudes from 
[Google](https://developers.google.com/public-data/docs/canonical/states_csv).

- I create voting percentages within the range of [0.4, 0.6] so that I could eventually color the map red and blue in a manner that closely reflected population political leanings.


```{r, eval=TRUE}
US2020_Election <- readxl::read_excel("Presidential_Results_Dec6_2020.xlsx") %>%
  mutate(Total = Trump+Biden) %>%
  mutate(Trump_Percent = 
           ifelse(Trump/Total>0.6, 0.6, ifelse(Trump/Total<0.4, 0.4, Trump/Total))) %>%
  mutate(Trump_Percent_All = Trump/Total)

US2020_Election %>% head()
```

```{r, eval=TRUE}
states %>% head()
```

- Merge datasets

```{r, eval=TRUE}
US2020_Election$region <- tolower(US2020_Election$State_Name) # small letters

mapdata <- merge(states, US2020_Election, by = "region")

View(mapdata)
```

- Draw a choropleth map by using "Trump_Percent" variable as the fill.

```{r, eval=TRUE}
ggplot(mapdata) +
  geom_polygon(aes(x=long,y=lat,group=group, fill=Trump_Percent)) +
  coord_map() + theme_bw()
```

- Change the color scheme.

```{r, eval=TRUE}
ggplot(mapdata) +
  geom_polygon(aes(x=long,y=lat,group=group, fill=Trump_Percent)) +
  coord_map() + theme_bw() +
  scale_fill_gradient2(low="blue", mid='purple', high="red", midpoint=0.5,
                          limits=c(.3,.7), name='Rep %')
```

```{r, eval=TRUE}
ggplot(mapdata) +
  geom_polygon(aes(x=long,y=lat,group=group, fill=Trump_Percent_All)) +
  coord_map() + theme_bw() +
  scale_fill_gradient2(low="blue", mid='purple', high="red", midpoint=0.5,
                          limits=c(.3,.7), name='Rep %')
```




# Package: `tmap`

- `tmap` is a powerful and flexible map-making package. It has a concise syntax that allows for the creation of attractive maps with minimal code which will be familiar to `ggplot2` users. 

- For more info, see [tmap-getstarted](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html), [Making maps with R](https://geocompr.robinlovelace.net/adv-map.html#fig:tmshape), as well as [Tennekes (2018)](https://www.jstatsoft.org/article/view/v084i06).

  
```{r}
library(sf)
library(tigris) # download shp file from U.S. census website
library(tmaptools)
library(tmap)

download.file("http://www2.census.gov/geo/tiger/GENZ2015/shp/cb_2015_us_state_20m.zip", destfile = "states.zip")
unzip("states.zip")
```

- Merge data, and delete the data. 

    - dplyr::inner_join(): includes all rows in x and y.

    - dplyr::left_join(): includes all rows in x.

    - dplyr::right_join(): includes all rows in y.

    - dplyr::full_join(): includes all rows in x or y.
    
    - To join by different variables on x and y, use a named vector. For example, by = c("a" = "b") will match x$a to y$b.

```{r, eval=TRUE}
usmap <- st_read("cb_2015_us_state_20m.shp", stringsAsFactors = FALSE)

usdata <- inner_join(usmap, US2020_Election, by=c("STUSPS" = "State_Abbreviation"))
str(usdata)
class(usdata)

usdata2 <- usdata %>% filter(!STUSPS %in% c("PR", "HI", "AK")) # delete Puerto Rico, Hawaii, and Alaska
```

- You can use `sf` class with `ggplot2`.

```{r, eval=TRUE}
ggplot(usdata2) + geom_sf(aes(fill=Trump_Percent)) + theme_bw() +
  scale_fill_gradient(low="beige", high="red", limits=c(.4,.6), name='Rep %')
```

- Let's use package `tmap`.

```{r, eval=TRUE}
tm_shape(usdata2) + tm_polygons(col = "Trump_Percent", id="STUSPS", palette = "Reds", title = "Rep %")
```
