---
title: "DS 2003: Introduction to Shiny: Part 2"
output:
     html_document:
          number_sections: yes
          toc: TRUE
          toc_depth: 4
          toc_float: true
          toc_collapsed: true
          theme: journal
          code_folding: hide
runtime: shiny          
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(foreign)
library(ggplot2)
library(shiny)
```

## Independent controls

- By default, input elements are independent of all the other input elements.

```{r}
ui <- fluidPage(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2),
  plotOutput("irisplot"))

server <- function(input, output){
    output$irisplot <- renderPlot({
      hist(iris$Sepal.Length, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Sepal Length", main = "Histogram and Density of Sepal Length from Iris Data")
  
      dens <- density(iris$Sepal.Length, adjust = input$bw_adjust)
      lines(dens, col = "blue")
    })}

shinyApp(ui = ui, server = server, options = list(height = 600)) 
```

## Interdependent controls

- But, some input elements may depend on other input elements.

- In such a case, you can (i) use `renderUI` to create an input element depending on server-side values and (ii) display the input element with `uiOutput`

- Suppose that the min of bandwidth adjustment is "0.2", and the max is "the number of bins /10". 

```{r}
ui <- fluidPage(
  selectInput("n_breaks2", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  uiOutput("bw_adjust_ui"),
  plotOutput("irisplot"))

server <- function(input, output){
  
  output$bw_adjust_ui <- renderUI({
      
      if(is.null(input$n_breaks2)){
        return()
      }
    
     sliderInput("bw_adjust2", label = "Bandwidth adjustment:",
              min = 0.2, max = as.numeric(input$n_breaks2)/10, value = (0.2 + as.numeric(input$n_breaks2)/10)/2, step = 0.2)
    })

    output$irisplot <- renderPlot({
      
     if(is.null(input$n_breaks2)){
        return()
      }  
      
      hist(iris$Sepal.Length, probability = TRUE, breaks = as.numeric(input$n_breaks2),
       xlab = "Sepal Length", main = "Histogram and Density of Sepal Length from Iris Data")
  
      dens <- density(iris$Sepal.Length, adjust = input$bw_adjust2)
      lines(dens, col = "blue")
    })}

shinyApp(ui = ui, server = server, options = list(height = 500)) 
```

## Common errors with interdependent controls

- If input elements are displayed before their values are set, it produces errors. To prevent the errors, add the following codes:
   
```
if(is.null(input$...)) {
  return()
}
```

# Reactive Expressions

- You can use *reactive* expressions to stop an expression being updated until you use a specified controller.

- *reactive* expressions are evaluated when *input$<id>* is updated. 

- *reactive* objects must be called as a function.

## without reactive expressions

```{r}
scs <- read.spss('SCS_QE.sav', to.data.frame=T)

ui <- fluidPage(
   sidebarLayout(
    sidebarPanel(   
    sliderInput("span", label = "Span adjustment:",
              min = 0.2, max = 0.9, value = 0.5, step = 0.1),
    sliderInput("sd", label = "Standard deviation of error:",
              min = 0.2, max = 2, value = 0.5, step = 0.2)  
    ),
  mainPanel(
     plotOutput("scsplot")
    )
  ))

server <- function(input, output){
   
   output$scsplot <- renderPlot({
     
     scs$mathpre2 <- scs$mathpre + rnorm(nrow(scs), mean = 0, sd = input$sd) # add noise
     
    ggplot(scs, aes(x=mars, y=mathpre2)) + geom_point(alpha=0.3) + geom_smooth(method="loess", formula=y~x, se=FALSE, span=input$span) + theme_bw() + ylim(-1, 20) + xlim(20, 130)
    })}

shinyApp(ui = ui, server = server) 
```

## with reactive expressions

- We want to re-evaluate the `mathpre2` with noise whenever *input$sd* is updated.

- Using reactive expressions it's possible to control when the `mathpre2` is invalidated.

```{r}
ui <- fluidPage(
   sidebarLayout(
    sidebarPanel(   
    sliderInput("span2", label = "Span adjustment:",
              min = 0.2, max = 0.9, value = 0.5, step = 0.1),
    sliderInput("sd2", label = "Standard deviation of error:",
              min = 0.2, max = 2, value = 0.5, step = 0.2)  
    ),
  mainPanel(
     plotOutput("scsplot")
    )
  ))

server <- function(input, output){
   
    scs_data <- reactive({
     scs$mathpre2 <- scs$mathpre + rnorm(nrow(scs), mean = 0, sd = input$sd2) # add noise
     scs
    })
      
   output$scsplot <- renderPlot({
      ggplot(scs_data(), aes(x=mars, y=mathpre2)) + geom_point(alpha=0.3) + geom_smooth(method="loess", formula=y~x, se=FALSE, span=input$span2)  + theme_bw() + ylim(-1, 20) + xlim(22, 128)
     })}

shinyApp(ui = ui, server = server) 
```

# About layouts

- `sidebarLayout`: A two column layout with `sidebarPanel` and `mainPanel` (inside a fluidpage)

    - `sidebarPanel` contains input controls.
    - `mainPanel` contains outputs. 

- `splitLayout`: equally-spaced multiple column layout (inside a fluidpage)
    
- `tabsetPanel`: tabsets that divide outputs into multiple independently viewable sections (inside a fluidpage)

- `navbarPage`: A horizontal navigation that creates multipage dashboards

- Note that `fluidpage` should be used to contain all layout elements except `navbarPage`.

- For more info, see [Shiny Cheatsheet](https://shiny.rstudio.com/images/shiny-cheatsheet.pdf).

##  `sidebarLayout()`

```
ui <- fluidPage(
      sidebarLayout( 
        sidebarPanel(),
           mainPanel() 
      )
    )  
```

```{r}
# example from IntroShiny1
ui <- fluidPage(
   sidebarLayout(
    sidebarPanel( 
     sliderInput("span_adj", label = "Span adjustment:",
              min = 0.2, max = 0.9, value = 0.5, step = 0.1),
  ),
    mainPanel(tabsetPanel(
      tabPanel("Scatterplot", plotOutput("irisplot")),
      tabPanel("Span Value", textOutput("span_val")),
      tabPanel("Data", dataTableOutput("data_table")),
      tabPanel("Variable Summary", verbatimTextOutput("summary_var")))
  )))

server <- function(input, output){
    
    output$irisplot <- renderPlot({
    ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + geom_point() +  
     geom_smooth(method="loess", formula=y~x, se=FALSE, span=input$span_adj) + theme_bw()
    })
    
    output$span_val <- renderText({
      paste0("Span = ", input$span_adj)
    })
    
    output$data_table <- renderDataTable({
      iris
    })
    
    output$summary_var <- renderPrint({
      summary(iris[, c("Sepal.Length", "Sepal.Width")])
    })
}

shinyApp(ui = ui, server = server) 
```

##  `splitLayout()` 

```
ui <- fluidPage( 
      splitLayout(# object 1,
                  # object 2
       )
      )
```

```{r}
ui <- fluidPage(
   splitLayout(
     sliderInput("span_adj2", label = "Span adjustment:",
              min = 0.2, max = 0.9, value = 0.5, step = 0.1),
    tabsetPanel(
      tabPanel("Scatterplot", plotOutput("irisplot")),
      tabPanel("Span Value", textOutput("span_val")),
      tabPanel("Data", dataTableOutput("data_table")),
      tabPanel("Variable Summary", verbatimTextOutput("summary_var")))
  ))

server <- function(input, output){
    
    output$irisplot <- renderPlot({
    ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + geom_point() +  
     geom_smooth(method="loess", formula=y~x, se=FALSE, span=input$span_adj2) + theme_bw()
    })
    
    output$span_val <- renderText({
      paste0("Span = ", input$span_adj2)
    })
    
    output$data_table <- renderDataTable({
      iris
    })
    
    output$summary_var <- renderPrint({
      summary(iris[, c("Sepal.Length", "Sepal.Width")])
    })
}

shinyApp(ui = ui, server = server) 
```

## `navbarPage()`

```
ui <- navbarPage(title = "Page", 
      tabPanel("tab 1", "contents"), 
      tabPanel("tab 2", "contents"), 
      tabPanel("tab 3", "contents"))
```

```{r}
ui <- navbarPage("navbarPage Example",
  tabPanel("IRIS data",
   splitLayout(
     sliderInput("span_adj3", label = "Span adjustment:",
              min = 0.2, max = 0.9, value = 0.5, step = 0.1),
     tabsetPanel(
      tabPanel("Scatterplot", plotOutput("irisplot")),
      tabPanel("Span Value", textOutput("span_val")),
      tabPanel("Data", dataTableOutput("data_table")),
      tabPanel("Variable Summary", verbatimTextOutput("summary_var")))
  )),
 tabPanel("SCS data",
   sidebarLayout(
    sidebarPanel(   
    sliderInput("span3", label = "Span adjustment:",
              min = 0.2, max = 0.9, value = 0.5, step = 0.1),
    sliderInput("sd3", label = "Standard deviation of error:",
              min = 0.2, max = 2, value = 0.5, step = 0.2)  
    ),
    mainPanel(
     plotOutput("scsplot")
    )
  ))        
)

server <- function(input, output){
    
    output$irisplot <- renderPlot({
    ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + geom_point() +  
     geom_smooth(method="loess", formula=y~x, se=FALSE, span=input$span_adj3) + theme_bw()
    })
    
    output$span_val <- renderText({
      paste0("Span = ", input$span_adj3)
    })
    
    output$data_table <- renderDataTable({
      iris
    })
    
    output$summary_var <- renderPrint({
      summary(iris[, c("Sepal.Length", "Sepal.Width")])
    })
    
    scs_data <- reactive({
     scs$mathpre2 <- scs$mathpre + rnorm(nrow(scs), mean = 0, sd = input$sd3) # add noise
     scs
    })
      
   output$scsplot <- renderPlot({
      ggplot(scs_data(), aes(x=mars, y=mathpre2)) + geom_point(alpha=0.3) + geom_smooth(method="loess", formula=y~x, se=FALSE, span=input$span3)  + theme_bw() + ylim(-1, 20) + xlim(22, 128)
     })
}

shinyApp(ui = ui, server = server, options = list(height = 500)) 
```

### `navbarPage()` with `navbarMenu()`

- Try with `collapsible = FALSE` instead of `collapsible = TRUE`.

```{r}
ui <- navbarPage("navbarPage Example",
  tabPanel("IRIS data",
   splitLayout(
     sliderInput("span_adj4", label = "Span adjustment:",
              min = 0.2, max = 0.9, value = 0.5, step = 0.1),
     tabsetPanel(
      tabPanel("Scatterplot", plotOutput("irisplot")),
      tabPanel("Span Value", textOutput("span_val")),
      tabPanel("Data", dataTableOutput("data_table")),
      tabPanel("Variable Summary", verbatimTextOutput("summary_var")))
  )),
  navbarMenu("SCS data",  
  tabPanel("Blue Line",
   sidebarLayout(
    sidebarPanel(   
    sliderInput("span4", label = "Span adjustment:",
              min = 0.2, max = 0.9, value = 0.5, step = 0.1),
    sliderInput("sd4", label = "Standard deviation of error:",
              min = 0.2, max = 2, value = 0.5, step = 0.2)  
    ),
    mainPanel(
     plotOutput("scsplot")
    )
   )),
  tabPanel("Red Line",
   sidebarLayout(
    sidebarPanel(   
    sliderInput("span5", label = "Span adjustment:",
              min = 0.2, max = 0.9, value = 0.5, step = 0.1),
    sliderInput("sd5", label = "Standard deviation of error:",
              min = 0.2, max = 2, value = 0.5, step = 0.2)  
    ),
    mainPanel(
     plotOutput("scsplot2")
    )
   ))
  ), collapsible = TRUE
)

server <- function(input, output){
    
    output$irisplot <- renderPlot({
    ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + geom_point() +  
     geom_smooth(method="loess", formula=y~x, se=FALSE, span=input$span_adj4) + theme_bw()
    })
    
    output$span_val <- renderText({
      paste0("Span = ", input$span_adj4)
    })
    
    output$data_table <- renderDataTable({
      iris
    })
    
    output$summary_var <- renderPrint({
      summary(iris[, c("Sepal.Length", "Sepal.Width")])
    })
    
    scs_data <- reactive({
     scs$mathpre2 <- scs$mathpre + rnorm(nrow(scs), mean = 0, sd = input$sd4) # add noise
     scs
    })
    
    scs_data2 <- reactive({
     scs$mathpre2 <- scs$mathpre + rnorm(nrow(scs), mean = 0, sd = input$sd5) # add noise
     scs
    })
      
   output$scsplot <- renderPlot({
      ggplot(scs_data(), aes(x=mars, y=mathpre2)) + geom_point(alpha=0.3) + geom_smooth(method="loess", formula=y~x, se=FALSE, span=input$span4)  + theme_bw() + ylim(-1, 20) + xlim(22, 128)
     })
   
  output$scsplot2 <- renderPlot({
      ggplot(scs_data2(), aes(x=mars, y=mathpre2)) + geom_point(alpha=0.3) + geom_smooth(method="loess", formula=y~x, se=FALSE, span=input$span5, col="red")  + theme_bw() + ylim(-1, 20) + xlim(22, 128)
     })
}

shinyApp(ui = ui, server = server, options = list(height = 600)) 
```


# shinythemes

- `shinythemes` package makes it easy to alter the overall appearance of your Shiny applications. 

- See options from https://rstudio.github.io/shinythemes/. 

- You need to include the following code to change the shinytheme:

```
fluidPage(theme = shinytheme("cerulean"),
  ...
)

navbarPage(theme = shinytheme("journal"),
  ...
)
```

```{r}
library(shinythemes)

ui <- navbarPage("navbarPage Example",
  theme = shinytheme("journal"),                    
  tabPanel("IRIS data",
   splitLayout(
     sliderInput("span_adj5", label = "Span adjustment:",
              min = 0.2, max = 0.9, value = 0.5, step = 0.1),
     tabsetPanel(
      tabPanel("Scatterplot", plotOutput("irisplot")),
      tabPanel("Span Value", textOutput("span_val")),
      tabPanel("Data", dataTableOutput("data_table")),
      tabPanel("Variable Summary", verbatimTextOutput("summary_var")))
  )),
  navbarMenu("SCS data",  
  tabPanel("Blue Line",
   sidebarLayout(
    sidebarPanel(   
    sliderInput("span6", label = "Span adjustment:",
              min = 0.2, max = 0.9, value = 0.5, step = 0.1),
    sliderInput("sd6", label = "Standard deviation of error:",
              min = 0.2, max = 2, value = 0.5, step = 0.2)  
    ),
    mainPanel(
     plotOutput("scsplot")
    )
   )),
  tabPanel("Red Line",
   sidebarLayout(
    sidebarPanel(   
    sliderInput("span7", label = "Span adjustment:",
              min = 0.2, max = 0.9, value = 0.5, step = 0.1),
    sliderInput("sd7", label = "Standard deviation of error:",
              min = 0.2, max = 2, value = 0.5, step = 0.2)  
    ),
    mainPanel(
     plotOutput("scsplot2")
    )
   ))
  ), collapsible = TRUE
)

server <- function(input, output){
    
    output$irisplot <- renderPlot({
    ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + geom_point() +  
     geom_smooth(method="loess", formula=y~x, se=FALSE, span=input$span_adj5) + theme_bw()
    })
    
    output$span_val <- renderText({
      paste0("Span = ", input$span_adj5)
    })
    
    output$data_table <- renderDataTable({
      iris
    })
    
    output$summary_var <- renderPrint({
      summary(iris[, c("Sepal.Length", "Sepal.Width")])
    })
    
    scs_data <- reactive({
     scs$mathpre2 <- scs$mathpre + rnorm(nrow(scs), mean = 0, sd = input$sd6) # add noise
     scs
    })
    
    scs_data2 <- reactive({
     scs$mathpre2 <- scs$mathpre + rnorm(nrow(scs), mean = 0, sd = input$sd7) # add noise
     scs
    })
      
   output$scsplot <- renderPlot({
      ggplot(scs_data(), aes(x=mars, y=mathpre2)) + geom_point(alpha=0.3) + geom_smooth(method="loess", formula=y~x, se=FALSE, span=input$span6)  + theme_bw() + ylim(-1, 20) + xlim(22, 128)
     })
   
  output$scsplot2 <- renderPlot({
      ggplot(scs_data2(), aes(x=mars, y=mathpre2)) + geom_point(alpha=0.3) + geom_smooth(method="loess", formula=y~x, se=FALSE, span=input$span7, col="red")  + theme_bw() + ylim(-1, 20) + xlim(22, 128)
     })
}

shinyApp(ui = ui, server = server, options = list(height = 600)) 
```
    
# self-contained vs split-file shinyApps 

- self-contained shinyApps
 
    - useful for small apps
    
    - can be easily added in R code chucks

- split-file shinyApps: `ui.R` & `server.R`

    - allow for easy separation of codes between the `ui.R` and `server.R` components.
    - can be included into R markdown presentations with `shinyAppDir()`. 
   
## Some tips for the ui.R and server.R

**< ui.R >** 

- Don't include input variables such as `input$<id>`

- Don't include output variables such as `output$<id>`

- Instead, please always quote the names of variables like `plotOutput("<id>")`

- If you like, change the size of the output chart in `plotOutput("<id>")`.

**< server.R >**

- Add a new content to display to the output object with the following code: 

```
    output$<id> <- renderPlot({ ...
````    

- Select the appropriate `render*` function for the output you're creating to send to the user side. 

- Use input values from the input objects using their names, such as `input$<id>`.


# shinyAppDir, iframe, and include_app

- `shinyAppDir()` and `<iframe ...>` can include your shiny apps in the presentation.        

## `shinyAppDir()`

- `shinyAppDir()` can include an entire Shiny application within an R Markdown document. 

- unzip `shinyApp.zip` (from Collab) that contains `ui.R`, `server.R` and `SCS_QE.csv` in your working directory.

```{r, eval=TRUE}
#shinyAppDir(appDir = "shinyApp", options = list(width = "100%", height = 550))
```

- You can embed a Shiny application located in another directory:

```{r}
#shinyAppDir(system.file("examples/06_tabsets", package = "shiny"), options = list(width = "100%", height = 550))
```

## `iframe`

- Get the URL for your shiny app that is uploaded on [ShinyApp.io](https://www.shinyapps.io). 

- Use the following code and change the URL. Change the sizes of the width and height.

<iframe src="https://youmi.shinyapps.io/shinyApp_mini/" width="100%" height="550"></iframe>

## `include_app()`

```{r test}
#knitr::include_app(url = "https://youmi.shinyapps.io/shinyApp_mini/", height="550")
```

# htmlwidgets in Shiny Apps

- htmlwidgets work just like R plots except they produce interactive web visualizations. Check available packages from http://www.htmlwidgets.org/index.html 

- Examples: [plotly](http://www.htmlwidgets.org/showcase_plotly.html), [highcharter](http://www.htmlwidgets.org/showcase_highcharts.html), etc

- If you use split files, load the packages that provide interactivity (i.e., `Plotly`) using `library()` in both the `ui.R` and `server.R` files.

## package `Plotly`

- Let's use [Plotly](https://plot.ly/r/) we discussed in class. Again, `plotly` allows you to easily convert your `ggplot2` graphics to an interactive web-based version. 

- Use `renderPlotly()` in the `server` argument, and use `plotlyOutput()` in the `ui` argument.

```{r}
library(plotly)

shinyApp(
  ui = fluidPage(
  sliderInput("span_p", label = "Span adjustment:",
              min = 0.2, max = 0.9, value = 0.5, step = 0.1),
  plotlyOutput("irisplot")
  ),
  
  server = function(input, output){
    output$irisplot <- renderPlotly({
      
    p <- ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + geom_point() + geom_smooth(method="loess", formula=y~x, se=FALSE, span=input$span_p) + theme_bw()
    
    ggplotly(p)
    
      })},
  options = list(height=600)
)    
```



