---
course: "DS2003: Communicating with Data"
class title: "Interactive Lecture: tidyr "
class number: "14"
---

```{r}
#import dependencies
library(tidyverse)
```
## Tidy Data
>“Happy families are all alike; every unhappy family is unhappy in its own way.” –– Leo Tolstoy

>“Tidy datasets are all alike, but every messy dataset is messy in its own way.” –– Hadley Wickham

Based on Hadley Wickham's [Tidy Data](https://vita.had.co.nz/papers/tidy-data.pdf) paper. 

Data is said to be tidy if it meets the following conditions:

* Each variable forms a column.
* Each observation forms a row (see wide versus long datasets)
  https://www.statology.org/long-vs-wide-data/
* Each type of observational unit forms a table.

Example:
```{r}
student_results <- read.csv('student_results.csv')
student_results
```

**Is this data tidy or untidy data?**

This data is untidy due to the following common issues of untidy data:

*Some column headers are values, not variable names*
* The column headers `term 1`, `term 2` and `term 3` are values. These need not to be column headers but values within another column like term.

*Multiple variables are stored in one column*
The column `sex_age` has two variables. For example Mike’s value for this column is m_12, meaning that he is a 12 year old male. The sex and age values should be in different columns.

*Variables are stored in both rows and columns*
The `test number` column is not a variable, it stores names of variables test 1 and test 2. Therefore, test 1 and test 2 need to be headers of columns as they are variables.

*Multiple types of observational units are stored in the same table*
The table above has data for two types of observational units: the student and student performance in two tests for each term. These different observational units need to be in their own tables for consistency. Therefore, a table for student should have the id, name, phone, sex and age. The table for performance should have the rest of the data, plus the id column which should be common in both tables.

**Let's tidy this data**

1. Turn 'term 1', 'term 2', 'term 3' into a column 'term'

* To do this we will use `pivot_longer` from the tidyr package

```{r}
student_results_clean <- student_results %>%
  pivot_longer(cols = c('term1', 'term2', 'term3'), names_to = 'term', values_to = 'score')

student_results_clean
```

2. Split 'sex_age' variable
* use `separate()` from tidyr

```{r}
student_results_clean <- student_results_clean %>%
  separate(sex_age, c('sex', 'age'), '_')

student_results_clean
```

3. Make 'test`' and 'test2' headers
* use pivot_wider() from tidyr

```{r}
student_results_clean <- student_results_clean %>%
  pivot_wider(names_from = test_number, values_from = score)

student_results_clean
```

4. Split the table into a 'student' table and 'test' table
* use dplyr functions

```{r}
student <- student_results_clean %>%
  select('id', 'name', 'phone', 'sex', 'age') %>% distinct()

student
```

```{r}
test <- student_results_clean %>%
  select('id', 'term', 'test1', 'test2')

test
```

## Students in-class lab:
* Import `dogs.csv'
* Review data and transform to tidy data using tidyverse methods

Import csv file
```{r}


```


Fill in values for patient, breed, address, guardian, for all rows (fill down same value to next value)
Hint: fill() expects NA for fills
```{r}

```


Pivot longer to combine height, length, width columns to dimension
```{r}

```


split weight between "weight" and "weight_units"
```{r}

```


Split visit_date to visit_num and date
```{r}

```


Separate previous dataframe into 
* guardian_df (guardian and address)
* dog_df (all other columns)
```{r}


```


Display only dog visits from dog_df that is the first visit (should have 3 rows)
```{r}


```


