---
title: "R Markdown Lesson"
author: "Natalie Kupperman"
course: 'DS 2003: Communicating with Data'
class: '1'
output:
  html_document: default
  pdf_document: default

---

The section at the top (between the `---`) is the YAML - this is the metadata for the document

R Markdown lets us generate reports that include our code and results. This is important for reproducibility, accountability, and learning. 

In R Studio, Markdown can be written in any of the white space. 

First, let's install the proper packages. Code is entered in a code chunk. 

```{r}
# dependencies
library(rmarkdown)
library(knitr)
```

R Markdown uses special characters to define text formats

```{r}
# Heading 1
## Heading 2
### Heading 3
#### Heading 4
##### Heading 5
###### Heading 6
```

# Heading 1
## Heading 2
### Heading 3
#### Heading 4
##### Heading 5
###### Heading 6

(When in a code chuck the # 'comments out' the following text)

```{r}
# *Italics*
# **Bold**
# ***Bold Italicized***
```


*Italics*
**Bold**
***Bold Italicized***

```{r}
# 1. Numbered
# 2. List
```

1. Numbered
2. List

```{r}
# - bullet lists
# * can be made with
# - hypens or
# * asterisks
```

- bullet lists
* can be made with
- hypens or
* asterisks

```{r}
# hyperlink
# [text to be hyperlinked](website)
```

[University of Virginia](virginia.edu)

```{r}
# > block text or quotation text
```

> a bit of block text or quotation text

```{r}
# `computer text`
```

`computer text`

----------------------------------- 


**More Resources**

* [R Markdown Reference Guide](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf?_ga=2.107746698.949398315.1659967887-326248043.1657566606)

* [Using R Markdown for Class](https://ntaback.github.io/UofT_STA130/Rmarkdownforclassreports.html)

* [R Markdown Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)

* [R Markdown Themes](https://www.datadreaming.org/post/r-markdown-theme-gallery/)

