---
course: "DS2003: Communicating with Data"
class title: "Descriptive Statistics in R"
class : "6"
---

# Descriptive Statistics

```{r}
# using mtcars dataset

head(mtcars)

```
### There are other well known packages for computing statistics in R (e.g., pysch).  For now we will stick to Base R

## Measures of Center

### MEAN
```{r}
# the mean for one variable
mean(mtcars$mpg)
```

```{r}
# the means for all columns in a data frame
colMeans(mtcars)
```

```{r}
# the means for all rows in a data frame
rowMeans(mtcars)
```

### MEDIAN
```{r}
median(mtcars$mpg)
```

### In base R there is no built in function to get the median of every column.
* We can use the apply() function 
* The apply() function takes a data object, a function to execute, and a specified margin (rows or columns)

```{r}
colMedians <- apply(mtcars,         # data object
                    MARGIN = 2,     # operate on columns (specified margin),
                    FUN = median)   # function to execute

colMedians
```

```{r}
# Students: use apply() to get the medians of all rows
colRows <- apply(mtcars,
                 MARGIN =1,
                 FUN = median)

colRows
```

## Comparing Mean and Median

### HISTOGRAM
* plot to show freqency distributions
* commonly used during data exploration
```{r}
hist(mtcars$mpg)
```

### DENSITY PLOT
* a density plot is a smoothed version of a histogram
* the y-axis is a probability density (i.e., the probability per unit on the x-axis)
* the total area under the curve integrates to 1
```{r}
# use plot()
plot(density(mtcars$mpg))

# plot a black line at the mean
abline(v = mean(mtcars$mpg, lwd = 5))

# plot a red line at the median
abline(v = median(mtcars$mpg), col = 'red', lwd = 2)
```
### Question: What type of skew is seen above?

```{r}
# combination of histogram and kernel density estimation
hist(mtcars$mpg, prob = T, xlab = 'mpg', main = 'MPG Distribution')
lines(density(mtcars$mpg), lwd = 3, col = 'red')
```


```{r}
data <- rnorm(50)            
otl <- rnorm(3, mean = 15)     
data <- c(data, otl)      
```

```{r}
# Students: Create a density plot with lines for median and mean for 'data' (generated above)

plot(density(data))                

abline(v=mean(data), lwd=5)        

abline(v=median(data), col="red", lwd=2 )
```
### Give 3 bullet points interpreting this plot
*
*
*

## MODE
### Can be used for categorical or numeric data
### Not horribly useful in statistics
```{r}
# with dpylr
mtcars %>% count(mpg, sort = TRUE)

```

```{r}
# with user defined function
mode_function <- function(data){                         # Define new function
    data_table <- table(data)                            # Create data table
    max_index <- which.max(data_table)                   # Find max index
    if (is.numeric(data)){                               # If input is numeric data
        return(as.numeric(names(data_table)[max_index])) # Return output as numeric
    }
    names(data_table)[max_index]            # Otherwise return output as character
}

mode_function(mtcars$mpg)
```


```{r}
# can place user defined function into apply()
colModes <- apply(mtcars,            
                 MARGIN=2,               
                 FUN = mode_function)    # use function mode_function

colModes
```

## Measures of Spread
## RANGE
```{r}
# Subtract min from max
max(mtcars$mpg) - min(mtcars$mpg)
```

```{r}
# range() gives min and max values
range(mtcars$mpg)
```

## QUANTILES
```{r}
# use quantile to get quartiles
quantile(mtcars$mpg)
```

## Other functions to get quartiles
```{r}
#five number summary
fivenum(mtcars$mpg)
```
```{r}
# Summary() shows the five number summary + the mean
summary(mtcars$mpg)
```

```{r}
quantile(mtcars$mpg,
        probs = c(0.1,0.9))  # get the 10th and 90th percentiles
```

```{r}
quantile(mtcars$mpg, 
         probs = seq(0, 1, .2)) # get quintiles using sequence function
```

## INTERQUARTILE RANGE (IQR)
```{r}
IQR(mtcars$mpg)
```

## BOXPLOT
```{r}
boxplot(mtcars$mpg)
```


```{r}
five_num <- fivenum(mtcars$mpg)

boxplot(mtcars$mpg)

text(x=five_num[1], adj=2, labels ="Minimum")
text(x=five_num[2], adj=2.3, labels ="1st Quartile")
text(x=five_num[3], adj=3, labels ="Median")
text(x=five_num[4], adj=2.3, labels ="3rd Quartile")
text(x=five_num[5], adj=2, labels ="Maximum")
text(x=five_num[3], adj=c(0.5,-8), labels ="IQR", srt=90, cex=2)
```

## VARIANCE
```{r}
# using var()
var(mtcars$mpg)
```

## STANDARD DEVIATION
```{r}
sd(mtcars$mpg)
```

## MEDIAN ABSOLUTE DEVIATION
### alternative measure of spread based based on the median
```{r}
mad(mtcars$mpg)
```

## DESCRIPTIVE STATISTICS ON SUBSETS

```{r}
mean(mtcars$mpg[mtcars$cyl == 4])  # mean for a subset
```


```{r}
# Students: Find the median mpg for cars with horsepower greater than or equal to 120
median(mtcars$mpg[mtcars$hp >= 120])
```

## Using tapply()
### Create group summaries based on factor levels
```{r}

tapply(mtcars$mpg,     # x
       mtcars$cyl,     # index (factor)
       mean)           # function
```


### tapply(incex$income, list(incex$edu, incex$occ), mean)  # by education & occ. status


## Use with() 
```{r}
# Alternatively use with() instead of repeated 'mtcars$"
with(mtcars, tapply(mpg, cyl, mean))
```

## Use attach() and detach()
```{r}
# makes all variable of the mtcars data frame available
attach(mtcars)    

head(income)

tapply(income, edu, mean) 

detach(incex)        # after you are done, detach data set

```

### !!! WARNING: in order to avoid confusions when manipulating data avoid using attach() and detach() !!! 
## But attaching data is useful for plotting   

## Basic Tables
### sapply() & apply()
```{r}
sapply(mtcars, mean)

sapply(mtcars, class)

sapply(mtcars, function(x) sum(is.na(x)))

apply(mtcars, 2, function(x) sum(is.na(x)))

apply(mtcars, 1, function(x) sum(is.na(x)))
```

## CROSSTABS
```{r}
xtabs(~ mpg + cyl, data = mtcars)  

xtabs(~ cyl, data = mtcars)

xtabs(mpg ~ cyl, data = mtcars) # sum of mpg by cyl

xtabs(mpg ~ cyl, data = mtcars) / xtabs( ~ cyl, data = mtcars) # mean mpg by cyl

with(mtcars, tapply(mpg, cyl, mean))

```

### Compare statistics
```{r}
with(mtcars, tapply(mpg, cyl, summary)) # summary

with(mtcars, tapply(mpg, cyl, mean))    # mean

with(mtcars, tapply(mpg, cyl, sd))      # standard deviation

library(tidyverse)
mtcars %>% count(cyl)
```

```{r}
# using boxplots

boxplot(mpg ~ cyl, data = mtcars, varwidth = F, horizontal = T) #default

boxplot(mpg ~ cyl, data = mtcars, varwidth = T, horizontal = T)

#Students : what is varwidth doing?
```








## FUNCTIONS

```{r}
# ::::: mean value & standard dev. :::::
my_mean <- function(x)
{
   # computes mean value of vector x
   sum(x) / length(x)
}

my_mean             # without (), prints the body of the function

my_mean(mtcars$mpg)


my_sd <- function(x)
{
   # computes standard deviation of vector x
   sqrt(sum((x - mean.val(x))^2) / (length(x) - 1))
}

my_sd(incex$income)

# define your own my.summary() function
my_summary <- function(x)
{
   # returns mean, sd, n, and standard error
   # x ... numeric vector (no missings allowed)

   m <- mean.val(x)
   s <- sd.val(x)
   n <- length(x)
   se <- s / sqrt(n)  # standard error
   c(mean = m, std.dev = s, n = n, s.e. = se)
}

my.summary(mtcars$mpg)
```


## OTHER PLOTS IN BASE R
### Barplot
```{r}
# barplot()
barplot(table(mtcars$cyl), main = 'cylinders')
```

```{r}
# plot()
plot(~ cyl, data = mtcars)      
```

### Scatter Plot
```{r}
plot(mpg ~ hp, data = mtcars)
```




