---
title: 'DS2003: Interactive Plots in plotly'
author: "Natalie Kupperman"
week: 9
class: 17
---
## Plotly

* `plotly` is a high-level interface to [plotly.js](https://github.com/plotly/plotly.js), based on [d3.js](https://d3js.org/) which provides an easy-to-use UI to generate slick D3 interactive graphics.

* Plot work in multiple formats: viewer windows, R Markdown documents, shiny apps

* plotly is also backed by a strong community and active development

* plotly works with R, python, excel, and others. We will just focus on R in this class.  

### Why use interactive graphics?
  - Zooming
  - Silencing
  - Hovering
  - Sliding
  - Filtering
  
### Difference between interactive and animated graphs
- interactive graphs allow the user to manipulate different features of the graph

  
### plotly Resources

- [The Plolty Website](https://plotly.com/)
- [The Plotly R documentation](https://plotly.com/r/)
- [The Plotly R Package on GitHub](https://github.com/plotly/plotly.R)
- [The Plotly R Cheatsheet](https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf)

```{r}
library(plyr)
library(tidyverse)
library(plotly)
```

```{r}
# read in 'wine.csv' data
wine <- read.csv('wine.csv')

head(wine)
```
```{r}
wine$Type <- as.factor(wine$Type)
```


## Convert Static Graphs to Interactive Graphs Using `ggplotly`

Let's look at a static graph comparing `flavanoids` to `proline`

```{r}
static_plot <- wine %>%
  ggplot(aes(Flavanoids, Proline, color = Type)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("#00AFBB", "#E7B800", "#FC4E07"))

static_plot
```

### How to create a plotly object

- There are two main approaches to initialize a plotly object:  
  
   + transforming a ggplot2 object with `ggplotly()` 
   + setting up aesthetics mappings with `plot_ly()` directly

## Today's focus `ggplotly()`

- `ggplotly()` takes existing `ggplot2` objects and converts them into interactive `plotly` graphics. That is, `ggplotly()` converts your static plots to an interactive web-based version!

- This makes it easy to create interactive figures because we are already familiar to the `ggplot2` syntax.

```{r}
# Create an interactive plot of Flavonoids vs Proline

ggplotly(static_plot)
```
### Notes
Interactive plots are great!  
*Bad Design = Bad Interactive Plots  
*Follow data-viz best practices  

### Bar Plot

```{r}
g <- ggplot(data = wine, aes(x = Type, fill = Type)) + 
     geom_bar(fill = c("#00AFBB", "#E7B800", "#FC4E07")) +
     theme_classic() 

ggplotly(g)
```

### If you want the bar in descending order
```{r}
g.5 <- count(wine, Type) %>% 
       ggplot(aes(x = reorder(Type, -n), y = n)) +
       geom_bar(fill = c("#00AFBB", "#E7B800", "#FC4E07"), stat = 'identity') +
       theme_classic() +
       xlab('Type')
     
ggplotly(g.5)
```

### Histogram

```{r}
g2 <- ggplot(wine, aes(x = Flavanoids, fill = as.factor(Type))) + 
      geom_histogram(bins = 15, alpha = 0.6) +
      theme_minimal() +
      scale_fill_manual(values = c("#00AFBB", "#E7B800", "#FC4E07")) 

ggplotly(g2)
```

### Scatterplot

```{r}
g4 <- ggplot(wine, aes(x = Flavanoids, y = Proline, color = as.factor(Type))) + 
      geom_point() + 
      geom_smooth(method="loess", formula=y~x, se=F) + 
      scale_color_manual(values = c("#00AFBB", "#E7B800", "#FC4E07"), name = 'Type') +
      theme_bw() 

ggplotly(g4)
```
```{r}
g5 <- ggplot(wine, aes(x = Flavanoids, y = Proline, color = as.factor(Type))) + 
      geom_point() + 
      geom_smooth(method="lm", formula=y~x, se=F) + 
      scale_color_manual(values = c("#00AFBB", "#E7B800", "#FC4E07"), name = 'Type') +
      theme_bw()

ggplotly(g5)
```

## Creating interactive graphs directly in `plot_ly`

- `plot_ly()` is the base `plotly` command to initialize a plot from a dataframe, similar to `ggplot()` from ggplot2.

### import `plotly`

```{r}
library(plotly)
```

### Bar plot
```{r}
wine %>% mutate(fType = as.factor(Type)) %>%
  count(fType) %>% #using dplyr create frequency table
  plot_ly(x = ~fType, y = ~n) %>% # initialize graph in plotly, use ~ for aes mappings
  add_bars() %>% #set graph as bar chart
  layout(title = "Freq. of Wine Soil Type",
         xaxis = list(title = "Type"),
         yaxis = list(title = "Frequency"))
```

### Reorder columns

```{r}
wine %>% mutate(fType = as.factor(Type)) %>%
  count(fType) %>% #using dplyr create frequency table
  mutate(fType = fct_reorder(fType, n, .desc = TRUE)) %>% #order for largest to smallest
  plot_ly(x = ~fType, y = ~n) %>% # initialize graph in plotly, use ~ for aes mappings
  add_bars() %>% #set graph as bar chart
  layout(title = "Freq. of Wine Soil Type",
         xaxis = list(title = "Type"),
         yaxis = list(title = "Frequency"))
```

### Basic Histogram
```{r}
g6 <- plot_ly(wine, x = ~Flavanoids, type = "histogram")

g6
```

### Flavanoids by Type Histogram

```{r}
Type1 <- wine %>% filter(Type == 1)
Type2 <- wine %>% filter(Type ==2)
Type3 <- wine %>% filter(Type == 3)

```

```{r}

g7 <- plot_ly(alpha = 0.4) %>%
  add_histogram(x = ~Type1$Flavanoids,
                name = 'Type 1',
                opacity = 0.6, 
                marker = list(color = "#00AFBB",
                              alpha = 0.6,
                              line = list(color = "lightgray",
                              width = 2))) %>%
  add_histogram(x = ~Type2$Flavanoids,
                name = 'Type 2',
                opacity = 0.6,
                marker = list(color = "#E7B800",
                              line = list(color = "lightgray",
                              width = 2))) %>%
  add_histogram(x = ~Type3$Flavanoids, 
                name = 'Type 3',
                opacity = 0.6,
                marker = list(color = "#FC4E07",
                              line = list(color = "lightgray",
                              width = 2))) %>%
  layout(barmode = 'overlay',
         title = 'Histogram of flavanoids by soil type',
         xaxis = list(title = 'flavanoids',
                      zeroline = FALSE),
         yaxis = list(title = 'count'))

g7

```
## Scatter plot

```{r}
fig <- plot_ly(type = 'scatter', mode = 'markers') 

fig <- fig %>%
  add_trace(
    x = Type1$Flavanoids,
    y = Type1$Proline,
    opacity = 0.5,
    marker = list(
      color = "#00AFBB",
      size = 7,
      line = list(
        color = 'lightgray',
        width = .5
      )
    ),
    name = 'Type 1'
  ) 

fig <- fig %>%
  add_trace(
    x = Type2$Flavanoids,
    y = Type2$Proline,
    marker = list(
      color = "#E7B800",
      size = 7,
      line = list(
        color = 'lightgray',
        width = .5
      )
    ),
    name = 'Type 2'
  )

fig <- fig %>%
  add_trace(
    x = Type3$Flavanoids,
    y = Type3$Proline,
    marker = list(
      color = "#FC4E07",
      size = 7,
      line = list(
        color = 'lightgray',
        width = .5
      )
    ),
    name = 'Type 3'
  )

fig <- fig %>%
  layout(title = 'Scatterplot of Flavanoids & Proline',
         xaxis = list(title = 'flavanoids',
                      zeroline = FALSE),
         yaxis = list(title = 'proline'))
fig
```

### 3D Scatterplot
```{r}
fig2 <- plot_ly(type = 'scatter3d', mode = 'markers') 
fig2 <- fig2 %>%
  add_trace(
    x = Type1$Flavanoids,
    y = Type1$Proline,
    z = Type1$Alcalinity,
    opacity = 0.5,
    marker = list(
      color = "#00AFBB",
      size = 7,
      line = list(
        color = 'lightgray',
        width = .5
      )
    ),
    name = 'Type 1'
  ) 

fig2 <- fig2 %>%
  add_trace(
    x = Type2$Flavanoids,
    y = Type2$Proline,
    z = Type2$Alcalinity,
    marker = list(
      color = "#E7B800",
      size = 7,
      line = list(
        color = 'lightgray',
        width = .5
      )
    ),
    name = 'Type 2'
  )

fig2 <- fig2 %>%
  add_trace(
    x = Type3$Flavanoids,
    y = Type3$Proline,
    z = Type3$Alcalinity,
    marker = list(
      color = "#FC4E07",
      size = 7,
      line = list(
        color = 'lightgray',
        width = .5
      )
    ),
    name = 'Type 3'
  )

axx <- list(
  title = "flavanoids"
)

axy <- list(
  title = "proline"
)

axz <- list(
  title = "alcalinity"
)

fig2 <- fig2 %>%
  layout(title = 'Scatterplot of Flavanoids, Proline & Alcalinity',
         scene = list(xaxis = axx,
                      yaxis = axy,
                      zaxis = axz))

fig2
```

## Line Plot
(From plotly examples)
```{r}

#create date
tg <- ddply(ToothGrowth, c("supp", "dose"), summarise, length=mean(len))


# create figure
fig <- plot_ly(tg, x = ~dose, y = ~length, type = 'scatter', mode = 'lines', linetype = ~supp, color = I('black')) 

fig <- fig %>% layout(title = 'The Effect of Vitamin C on Tooth Growth in Guinea Pigs by Supplement Type',
         xaxis = list(title = 'Dose in milligrams/day'),
         yaxis = list (title = 'Tooth length'))

fig
```

