---
course: 'DS2003: Communicating with Data'
class title: 'Introduction to R: Part 2'
class number: '3'
---

#################################
# Matrix
#################################
### In R, a matrix is a collection of elements of the same data type (numeric, character, or logical) arranged into a fixed number of rows and columns

### Since you are working with rows and columns, a matrix is two-dimensional

### You can construct a matrix in R with the matrix() function

```{r}
# Example
matrix(1:12, byrow = TRUE, nrow = 2)
```

In the matrix() function:

* The first argument is the collection of elements that R will arrange into the rows and columns of the matrix. Here, we use 1:12 which is a shortcut for c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)

* The argument byrow indicates that the matrix is filled by the rows. 

* If we want the matrix to be filled by the columns, we just place byrow = FALSE

* The third argument nrow indicates that the matrix should have three rows

```{r}
# (s2) Make a matrix of the box office earnings for the movie The Matrix
# the first element is US box offices, the second element is non-US box offices (in millions)
matrix_mv <- c(171.479, 465.974 )
reloaded <- c(281.576, 738.576)
revolutions <- c(139.313, 427.300 )
resurrections <- c(37.686, 156.467)

# Use c() to combine the three vectors into one vector called this vector box_office

box_office <- c(matrix_mv, reloaded, revolutions, resurrections)
box_office

```

```{r}
# (s3) Construct a matrix with 4 rows, where each row represents a movie. Call this matrix matrix_matrix

matrix_matrix <- matrix(box_office, byrow = TRUE, nrow = 4)
matrix_matrix

```

### Similar to vectors, you can add names for the rows and the columns of a matrix

* rownames(my_matrix) <- row_names_vector
* colnames(my_matrix) <- col_names_vector

```{r}
# Example: create vectors of row and column names

# rows
my_title <- c('Matrix', 'Matrix: Reloaded', 'Matrix: Revolutions', 'Matrix: Resurrections')
my_title

# columns 
my_region <- c('US', 'Non-US')
my_region

```

```{r}
# (s4) Use colnames() to name the columns of matrix_matrix with the region vector

colnames(matrix_matrix) <- my_region

# Use rownames() to the name the columns of the matrix_matrix with the title vector

rownames(matrix_matrix) <- my_title

# print the matrix_matrix

matrix_matrix
```

### rowSums() adds across columns and creates a new vector

```{r}
# (s5) Sum the rows of the matrix_matrix and save to a vector called worldwide

worldwide <- rowSums(matrix_matrix)
worldwide

```

### The above creates a vector but does not add it to the matrix
### We can do this using cbind()
### cbind() merges matrices and/or vectors together by column (big_matrix <- cbind(matrix1, matrix2, vector1)

```{r}
# (s6) merge matrix_matrix and worldwide using cbind() and call it all_matrix
all_matrix <- cbind(matrix_matrix, worldwide)

all_matrix
```

### rbind() does the same action, just with rows
```{r}
# Example: create another matrix with Star Wars data

box_office <- c(460.998, 314.4, 290.475, 247.900, 309.306, 165.8)
region <- c("US", "non-US")
titles <- c("A New Hope", 
            "The Empire Strikes Back", 
            "Return of the Jedi")

star_wars_matrix <- matrix(box_office, 
                           nrow = 3, byrow = TRUE,
                           dimnames = list(titles, region))

star_wars_matrix

```

```{r}
# (s7) use rbind() to merge star_wars_matrix and matrix_matrix

movies_matrix <-rbind(matrix_matrix, star_wars_matrix)
movies_matrix

```

### colSums() is similar rowSums() but sums down columns instead of row.

```{r}
# (s8) use colSums() to sum the box office revenue for the Matrix and Star Wars movies for the US and non-US 

colSums(movies_matrix)

```

### Matrix Selection
Similar to vectors, you can use the square brackets `[ ]` to select one or multiple elements from a matrix. Whereas vectors have one dimension, matrices have two dimensions. You should therefore use a comma to separate the rows you want to select from the columns. For example:

*my_matrix[1,2] selects the element at the first row and second column
*my_matrix[1:3,2:4] results in a matrix with the data on the rows 1, 2, 3 and columns 2, 3, 4

If you want to select all elements of a row or a column, no number is needed before or after the comma, respectively:

*my_matrix[,1] selects all elements of the first column
*my_matrix[1,] selects all elements of the first row

```{r}
# (s9) Use mean() to calculate the mean box office revenue for the US for the Matrix movies

mean(matrix_matrix[ ,1])

```

### You can use the arithmetic operators on a matrix
### 3 * matrix with multiple each element by 3

```{r}
# (s10) divide the matrix_matrix by 5

matrix_matrix/5

```

### you can also do matrix math where matrix * matrix multiples the element in the first matrix by the corresponding element is the second matrix - the result is a matrix of the products

```{r}
# Example
matrix1 <- matrix(1:10, nrow = 5)
matrix2 <- matrix(11:20, nrow = 5)

matrix1 * matrix2

```

###############################
# Factors
###############################

### Factor refers to a statistical data type used to store categorical variables

### The difference between a categorical variable and a continuous variable is that a categorical variable can belong to a limited number of categories. A continuous variable, on the other hand, can correspond to an infinite number of values

* It is important that R knows whether it is dealing with a continuous or a categorical variable, as the statistical models you will develop in the future treat both types differently

### Biological sex is good example of a categorical variable. In many circumstances you can limit the sex categories to "Male" or "Female"

* Capturing a person's gender requires more options but you will always have a finite number of categories

### R uses factor() to assign the factor data type to elements

```{r}
# Example

# create a vector of biological sex elements with 2 levels: Male & Female
biological_vector <- c('Male', 'Female', 'Female', 'Male', 'Female', 'Male', 'Male')

# assign biological_vector to a factor & assign to the variable factor_biological_vector
factor_biological_vector <- factor(biological_vector)
biological_vector

```

### Nominal categorical variable: categorical variable without an implied order
* cannot use logical operators to compare levels

### Ordinal categorical variable: cetegorical variable  with a natural order (e.g. high, medium, low)
* factor(ordinal_vector, order = TRUE, levels = c("level 1", "level 2", "level 3"))
* can use logical operators to compare levels

```{r}
# (s11) Create a factor vector using nominal variable and assign it to nominal

nominal <- c('blue', 'orange', 'white')
factor(nominal)
nominal

# Create a factor vector using ordinal variable and assign it to ordinal (Hint: see argument format above)

ordinal <- c('good', 'better', 'best')
factor(ordinal, order = TRUE, levels = c('good', 'better', 'best'))


```

# You can change the name of the levels by using the levels() 
* if you don't specify the order, R will assign the order alphabetically

```{r}
# Example: change the factor_biological_vector levels to M for male and F for female

# Check order of levels
factor_biological_vector

# change levels
levels(factor_biological_vector) <- c("F", "M")

# print vector
factor_biological_vector
```

```{r}
# (s12) Try out the summary() function on factor_biological_vector; use ? if you need help

summary(factor_biological_vector)

```

################################
# Lists
################################

### A list in R allows you to gather a variety of objects under one name (that is, the name of the list) in an ordered way. These objects can be matrices, vectors, data frames, even other lists, etc

### It is not even required that these objects are related to each other in any way

### list()

```{r}
# Example: create different data structures

# Vector 
my_vector <- 1:10 

# Matrix 
my_matrix <- matrix(1:9, ncol = 3)

# First 10 elements of the built-in data frame mtcars 
# We will get to dataframes in the next lesson
df <- mtcars[1:10,]

# Put together in a list
my_list <- list(my_vector, my_matrix, df)
my_list[[1]]
my_list[[2]]
```

```{r}
# Can give names to components, just like in vectos using name()

my_list <- list(my_vector, my_matrix, df)

names(my_list) <- c('v', 'm', 'd')

# print
my_list
my_list['v']
my_list['m']
my_list['d']

my_list[[1]]
my_list[[2]]
my_list[[3]]

```

