---
title: "Introduction to Shiny: Part 1"
class: "DS2003: Communicating with Data"
output:
     html_document:
          number_sections: yes
          toc: TRUE
          toc_depth: 4
          toc_float: true
          toc_collapsed: true
          theme: journal
          code_folding: show
runtime: shiny          
---

## Deploy Presentations to shinyapps.io

- Include "runtime: shiny" in the preamble. 

```
---
title: "Example"
output: ioslides_presentation
runtime: shiny
---
```
- Click "Run Presentation"
- Click "Publish" in the top-right corner of the shiny app.




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(shiny)
```



# What is Shiny?

- Framework for building interactive applications using the R programming language

    - Any R script or visualization can be made interactive using Shiny.

- An R package that makes it easy to build interactive web apps straight from R


- You can host standalone apps on a webpage or embed them in R Markdown documents or build dashboards.

- There is no need to learn HTML or JavaScript. 



# The basic structure
- Standard Application: GUIs <-- requests/callbacks -->  source code

- Web Application: Client GUI  <-- requests/callbacks --> server

- requests/callbacks can be event



# Run your Shiny app

- You can run through either your local machine or through the web browser, [ShinyApp.io](https://www.shinyapps.io).

- Local machine needs Rstudio and R package `shiny`.

- You need to use [ShinyApps.io](https://www.shinyapps.io) if you want to share you Shiny app with other people through the web browser.

   - [ShinyApp.io](https://www.shinyapps.io) is a hosting platform for Shiny apps.
   
   - Review [ShinyApps.IO Pricing](https://www.shinyapps.io/#pricing)
   
        - While your shiny apps are displayed on screen, they use active hours.

        - A default timeout period is 15 mins.
        
        
      
## Connect RStudio to your ShinyApps.io account

- Register for an account at [ShinyApps.io](https://www.shinyapps.io) 

- Get 25 active hours with a free shinyapps.io account

- Connect RStudio to your shinyapps.io account through the *Publishing Account* section of the *Global Options* under *Tools* (i.e.,  Tools > Global Options > Publishing Account). 



## Manage your deployed Shiny Apps

- You can manage your deployed Shiny apps at [ShinyApps.io](https://www.shinyapps.io).

- You can track usage activities (i.e., engagement and remaining active hours) through the dashboard at [ShinyApps.io](https://www.shinyapps.io)

- You can archive your shiny apps and then delete them if necessary.



# Embedded Shiny Application

- In Shiny App, there are two main arguments: `ui` and `server`.

- The `ui` defines what we see inside the ShinyApp and contains ''user/client-side code''.

     - A fluidpage is required for our inputs and outputs to be displayed together.
     
        - Input objects
        
        - `*output` functions are used to display output objects in the client side. 

- The `server` defines what we do depending on both inputs and outputs.

     - `render*` functions are used to generate output objects and depend on `input$<id>`.
     
     - output$[tag] <- [expression] updates the output object on the client side

- To display content from the `server` argument in the `ui`, you should use a pair of `render_*` and `*_Output` functions. For more info, see [Shiny Cheatsheet](https://shiny.rstudio.com/images/shiny-cheatsheet.pdf).

    - There are unique pairings of `render_*` and `*_Output` functions for all content types. 
    
    - Below, we focus on `renderPlot` in the `server` and `plotOutput` in the `ui`.
    
    
    
    
    

## Control Type: `sliderInput` 

- For demonstrations, we use here `sliderInput` function to create the input widgets used to drive an interactive plot. 
- For more info, see [Widget gallery](https://shiny.rstudio.com/gallery/widget-gallery.html).

```{r}

ui = fluidPage(
  # Input Objects
  sliderInput(inputId = "span", label = "Span adjustment",
              min = 0.2, max = 0.9, value = 0.5, step = 0.1),
  sliderInput("span0", label = "Subset data",
              min = 1, max = nrow(iris), value = nrow(iris), step = 10),
  # Output Objects
  plotOutput(outputId = "irisplot")
)

server = function(input, output) {
  

  # Update function that pairs with plotOutput: irisplot
  output$irisplot <- renderPlot({
    browser()
    my_iris<-iris[1:input$span0,]
    # my_iris<-iris 

    ggplot(my_iris, aes(x=Sepal.Length, y=Sepal.Width)) + geom_point() + 
      geom_smooth(method="loess", formula=y~x, se=FALSE, span=input$span) + theme_bw()
  })}

shinyApp(ui, server)


```


## Example: if you want to change the title depending on your input

```{r, include=TRUE}

ui = fluidPage(
  # Input Objects
  sliderInput(inputId = "span", label = "Span adjustment",
              min = 0.2, max = 0.9, value = 0.5, step = 0.1),
  # output Objects
  plotOutput(outputId = "irisplot")
)

server = function(input, output) {
  
  # Update function that pairs with plotOutput: irisplot
  output$irisplot <- renderPlot({
    my_iris<-iris
    ggplot(my_iris, aes(x=Sepal.Length, y=Sepal.Width)) + geom_point() + 
      geom_smooth(method="loess", formula=y~x, se=FALSE, span=input$span) + 
      ggtitle(paste0("Span: ", input$span)) +
      theme_bw()
  })}

shinyApp(ui, server)

```

## Control Type: `selectInput`

- We modify the above plot using `selectInput`. 

- The loess line is not added to the scatterplot. Why?

```{r}

ui = fluidPage(
  # Input Objects
    selectInput("span2", label = "Span adjustment:",
              choices = seq(0.2, 0.9, 0.1), 
              selected = 0.5),
  # output Objects
  plotOutput(outputId = "irisplot")
)

server = function(input, output) {
  
  # Update function that pairs with plotOutput: irisplot
  output$irisplot <- renderPlot({
    ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + geom_point() + 
      geom_smooth(method="loess", formula=y~x, se=FALSE, span=as.numeric(input$span2)) + 
      theme_bw()
  })}

shinyApp(ui, server)


```


### Specifying text for each value

```{r, include=TRUE}


ui = fluidPage(
  # Input Objects
      selectInput("span3", label = "Span adjustment:",
              choices = list("Span = 0.2" = 0.2, "Span = 0.3" = 0.3, "Span = 0.4" = 0.4,
                             "Span = 0.5" = 0.5, "Span = 0.6" = 0.6, "Span = 0.7" = 0.7,
                             "Span = 0.8" = 0.8, "Span = 0.9" = 0.9), 
              selected = 0.5),
  # output Objects
  plotOutput(outputId = "irisplot")
)

server = function(input, output) {
  
  # Update function that pairs with plotOutput: irisplot
  output$irisplot <- renderPlot({
    # print(input$span3)
    ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + geom_point() + 
      geom_smooth(method="loess", formula=y~x, se=FALSE, span=as.numeric(input$span3)) + 
      theme_bw()
  })}

shinyApp(ui, server)


```

## Notes

- `selectInput` produces strings, so if necessary, you need to convert them into other data types, say using `as.numeric`. 

- `print(input$<id>)` and `browser()` can be used in `render*` functions for debugging and troubleshooting.




## Using multiple `render*` and `*Output` functions

```{r}
shinyApp(
  ui = fluidPage(
  sliderInput("span5", label = "Span adjustment:",
              min = 0.2, max = 0.9, value = 0.5, step = 0.1),
  
  plotOutput("irisplot"), # add plot
  
  textOutput("span_val"), # add text
  
  dataTableOutput("data_table"), ## add data table
    
  verbatimTextOutput("summary_var") # add print output
  
  ),
  
  server = function(input, output){
    
    output$irisplot <- renderPlot({
    ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + geom_point() +  
     geom_smooth(method="loess", formula=y~x, se=FALSE, span=input$span5) + theme_bw()
    })
    
    output$span_val <- renderText({
      paste0("Span = ", input$span5)
    })
    
    output$data_table <- renderDataTable({
      iris
    })
    
    output$summary_var <- renderPrint({
      summary(iris[, c("Sepal.Length", "Sepal.Width")])
    })
  }                                  
)
```


## Changing layouts

### `SidebarLayout()`

- Use `SidebarLayout()` to display our content. 

```
ui <- fluidPage(
      sidebarLayout( 
        sidebarPanel(),
           mainPanel() 
      )
    )  
```
 
```{r}

ui = fluidPage(
  sidebarLayout(
    sidebarPanel( 
      sliderInput("span6", label = "Span adjustment:",
                  min = 0.2, max = 0.9, value = 0.5, step = 0.1),
    ),
    mainPanel(
      plotOutput("irisplot"),
      textOutput("span_val"),
      dataTableOutput("data_table"),
      verbatimTextOutput("summary_var")
    )
  )
)

server = function(input, output){
  
  output$irisplot <- renderPlot({
    ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + geom_point() +  
      geom_smooth(method="loess", formula=y~x, se=FALSE, span=input$span6) + theme_bw()
  })
  
  output$span_val <- renderText({
    paste0("Span = ", input$span6)
  })
  
  output$data_table <- renderDataTable({
    iris
  })
  
  output$summary_var <- renderPrint({
    summary(iris[, c("Sepal.Length", "Sepal.Width")])
  })
}                                  

shinyApp(ui, server)
```


### `tabsetPanel()`

- Additionally, use `tabsetPanel()` in the `mainPanel()` to display our content. 

```
ui <- fluidPage( tabsetPanel( 
  tabPanel("tab 1", "contents"), 
  tabPanel("tab 2", "contents"), 
  tabPanel("tab 3", "contents")))
```


```{r}
shinyApp(
  ui = fluidPage(
   sidebarLayout(
    sidebarPanel( 
     sliderInput("span7", label = "Span adjustment:",
              min = 0.2, max = 0.9, value = 0.5, step = 0.1),
  ),
    mainPanel(tabsetPanel(
      tabPanel("Scatterplot", plotOutput("irisplot")),
      tabPanel("Span Value", textOutput("span_val")),
      tabPanel("Data", dataTableOutput("data_table")),
      tabPanel("Variable Summary", verbatimTextOutput("summary_var")))
    )
   )
  ),
  
  server = function(input, output){
    
    output$irisplot <- renderPlot({
    ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + geom_point() +  
     geom_smooth(method="loess", formula=y~x, se=FALSE, span=input$span7) + theme_bw()
    })
    
    output$span_val <- renderText({
      paste0("Span = ", input$span7)
    })
    
    output$data_table <- renderDataTable({
      iris
    })
    
    output$summary_var <- renderPrint({
      summary(iris[, c("Sepal.Length", "Sepal.Width")])
    })
  }                                  
)
```

# Shiny Resources

- [Shiny](https://shiny.rstudio.com)!
- [Shiny Cheatsheet](https://shiny.rstudio.com/images/shiny-cheatsheet.pdf)
- [Shiny Documents](https://bookdown.org/yihui/rmarkdown/shiny-documents.html)
- [1st Shiny Competition](https://blog.rstudio.com/2019/04/05/first-shiny-contest-winners/),
- [Gallery](https://shiny.rstudio.com/gallery/)
- [Get Started](https://shiny.rstudio.com/tutorial/)
  - [First Shiny App](https://shiny.rstudio.com/tutorial/written-tutorial/lesson1/) 
    - which is discussed in this document
- ["Curly Brackets" in Shiny](https://community.rstudio.com/t/help-understanding-shiny-function-syntax-expression-objects-and-curly-brackets/22807/2)



