---
course: "DS2003: Communicating with Data"
class title: "Interactive Lecture: Beyond ggplot2 "
class number: "10"
---

## Beyond ggplot2

- ggplot2 is huge! About 50 geoms, 25 stats, 60 scales.

- Many extensions are very niche specific and developed by experts in the field.

- 115 registered extensions available to explore (https://exts.ggplot2.tidyverse.org)

- e.g., for plot compositions, you might want to use `gridExtra::grid.arrange()`, `ggpubr::ggarrange()`, `cowplot::plot_grid()`, or `patchwork`.

```{r}
# import libraries
library(tidyverse)
```

```{r}
# import data

df <- read.csv('data.csv')
```

## Examples

### Package: gridExtra

```{r, message=F, out.width='90%', fig.height=3.5, fig.align="center"}
library(gridExtra)

p1 <- ggplot(df) + geom_bar(aes(x = occ, fill=sex)) # barplot

p2 <- ggplot(df) + geom_histogram(aes(x=income), bins=10) # histogram

p3 <- ggplot(df) + geom_point(aes(x = age, y=income)) # scatterplot

grid.arrange(p1, p2, p3, nrow=1)
```

```{r}
grid.arrange(p3, arrangeGrob(p1, p2, ncol=2), nrow = 2)
```

### Package: patchwork

```{r, out.width='90%', fig.height=3.6, fig.align="center"}
library(patchwork)
p1 + p2 + p3

```

```{r}
( p1 + p2 ) / p3
```


### Package: GGally

```{r, message=F}
library(GGally)
ggpairs(df[, c('income', 'oexp', 'age', 'edu')])
```

```{r}
# 2 continous variables
ggpairs(df, columns = c(6,2), ggplot2::aes(colour = edu)) 
```

```{r}
# 1 continuous, 1 categorical variable
ggpairs(df, columns = c(6,1), ggplot2::aes(colour = edu)) 
```

```{r}
#heatmap
ggcorr(df)
```

```{r}
ggpairs(
  df[, c(6, 1, 3, 2)],
  upper = list(continuous = "density", combo = "box_no_facet"),
  lower = list(continuous = "points", combo = "dot_no_facet")
)
```

## Package: corrgram
```{r}
library(corrgram)
corrgram(df, 
         order=TRUE, 
         lower.panel=panel.ellipse, 
         upper.panel=panel.pts, 
         text.panel=panel.txt, 
         diag.panel=panel.minmax, 
         main="Income Survey") 

```

