---
course: 'DS2003: Communicating with Data'
class title: 'Introduction to R: Part 3'
class number: '4'
output:
  pdf_document: default
  html_document: default
---

################################
# Data frames (!!!)
################################

### In matrices we can only have one data type. In a data frame we can have a mix of data types
### As a data scientist, you will work a lot with data frames

```{r}
# Example: Often data frames will be imported via .csv, however, for learning, R has some pre-loaded datasets
# We will load the mtcars data frame to explore

# print mtcars data frame
mtcars

```
### Useful functions
```{r}
# head(): prints first 6 rows
head(mtcars)
```

```{r}
# tail(): prints last 6 rows of data
tail(mtcars)
```

```{r}
# str(): shows structure of dataset
str(mtcars)
```

```{r}
# dataframe(): creates a dataframe from vectors

# Definition of vectors
name <- c("Mercury", "Venus", "Earth", 
          "Mars", "Jupiter", "Saturn", 
          "Uranus", "Neptune")

type <- c("Terrestrial planet", 
          "Terrestrial planet", 
          "Terrestrial planet", 
          "Terrestrial planet", "Gas giant", 
          "Gas giant", "Gas giant", "Gas giant")

diameter <- c(0.382, 0.949, 1, 0.532, 
              11.209, 9.449, 4.007, 3.883)

rotation <- c(58.64, -243.02, 1, 1.03, 
              0.41, 0.43, -0.72, 0.67)

rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)

planets_df <- data.frame(name, type, diameter, rotation, rings)

planets_df
```

## Selection of data frame elements
### Similar to vectors and matrices, you select elements from a data frame with the help of square brackets [ ]
### by using a comma, you can indicate what to select from the rows and the columns respectively

```{r}
# Example: select columns 2-4 and rows 5-10 from mtcars

mtcars[5:10, 2:4]
```

```{r}
# Example: From mtcars subset columns 1 & 4 with all rows
mtcars[ , c(1, 4)]
```

```{r}
# Example: subset rows 20-25 & column 'mpg' of mtcars using the columns name

mtcars[20:25, 'mpg']
```

```{r}
# Example

mtcars[ , 1]

# =

mtcars[ , 'mpg']
```

```{r}
# select mpg and wt where cyl where cyl equals 8

mtcars[cyl == 8, c('mpg', 'wt')]  

```


```{r}
# Example: using the $ symbol

mtcars$mpg

```

```{r}
# Example: subset() - returns all entries whose mpg is greater than 10
# Can save the subset to a new df

mpg_10 <- subset(mtcars, mpg > 10)
mpg_10

```

```{r}
# Example: Returns entries where mpg is greater than 10 and cylinders are equal to 6

subset(mtcars, c(mpg > 10 & cyl == 6))

```

```{r}
# Example: Returns the columns cyl, mpg, hp for entries where cyl equals 4 and mpg is greater than 20

subset(mtcars, c(cyl == 4, mpg > 20), c(cyl, mpg, hp))

```
```{r}
# Misc
# get dimensions of df
dim(mtcars)       

# only one argument in brackets since a vector has only one dimension (rows)
length(mtcars$mpg)

# select each 5th value (starting with first obs.)                   
mtcars$mpg[seq(1, length(mtcars$mpg), by = 5)]  

# all values except values #10 to #20
mtcars$mpg[-(10:20)]
```


```{r}
# Example: sort() - works on vectors
# goes smallest -> largest

sort(mtcars$mpg)

```

```{r}
# Example: order() returns the sorting index

order(mtcars$mpg)

```


```{r}
# Example: order entries by mpg within dataframe

position <- order(mtcars$mpg)

mtcars[position, ]
```

################################
# Loading in data
################################

```{r}
# find out current working directory
getwd()

# set working directory to my laptop desktop
setwd('/Users/nataliekupperman/Desktop') # working directory on my laptop

# set working directory back to One Drive
setwd("/Users/nataliekupperman/Library/CloudStorage/OneDrive-UniversityofVirginia/Courses/DS2003/DS-3003/01 Intro to Course & R/class 2") 
```

## TABULAR separated data
### read in data with read.table() & sep = "\t
```{r}
# read in data
incex <- read.table('../data/income_exmpl.dat', header = T, sep = "\t") # use (..) to move up a directory

# print the first 6 lines of data with function head()
head(incex)       
```

### if the data are not located in the working directory (see setwd()) then you can read the data by specifying the full path

```{r}
# Example
incex <- read.table('/Users/nataliekupperman/Library/CloudStorage/OneDrive-UniversityofVirginia/Courses/DS2003/DS-3003/01 Intro to Course & R/data/income_exmpl.dat', header = T, sep = "\t")
```

### if an R function requires a character string or name as an argument (e.g., the file name or path), it must be set in quotes ("name" or 'name')

##COMMA separated data
### use read.csv()
```{r}
incex.2 <- read.csv('../data/income_exmpl.csv')
```

### Read other data formats (e.g., SPSS, SAS, or STATA)
*use the 'foreign'-package 
 
```{r}
#install.packages('foreign', repos = 'http://cran.cnr.Berkeley.edu/') 

# load package (makes its functions available)
library(foreign) 

# help on the package
library(help = 'foreign')  

# help on reading SPSS data files (.sav)
?read.spss  

incex.3 <- read.spss('..data/income_exmpl.sav', to.data.frame = TRUE)
```


