---
course: 'DS2003: Communicating with Data'
class title: 'Introduction to R: Part 1'
class number: '2'
output:
  pdf_document: default
  html_document: default
---
##############################
# R as a simple calculator
##############################

### Arithmetic operators
* Addition: +
* Subtraction: -
* Multiplication: *
* Division: /
* Exponentiation: ^ (e.g. 4^4 = 16)
* Modulo: %% (returns the remainder of the division of the number to the left by the number on its right; e.g. 5 %% 3 = 2)
* Square root: sqrt() (this a function vs an operator; e.g. sqrt(16) = 4)


```{r}
# Example: Addition

2 + 1

```

```{r}
# (s1) Addition - write an addition equation

```

```{r}
# (s2) Subtraction - write a subtraction equation

```

```{r}
# (s3) Multiplication - write a multiplication equation

```

```{r}
# (s4) Division - write a division equation

```

```{r}
# (s5) Exponentiation - write an exponentiation equation

```

```{r}
# (s6) Modulo - write a modulo equation

```

### Remember PEMDAS (Parentheses, Exponents, Multiplication, Division, Addition, Subtraction)

```{r}
# Equation with PEMDAS

100/(27-2)

#versus

100/27-2
```

```{r}
# (s7) Write an equation where PEMDAS is applicable

```

#################################
# Logical (comparison) operators
#################################

* < for less than
* > for greater than
* <= for less than or equal to
* >= for greater than or equal to
* == for equal to each other
* != not equal to each other

```{r}
# Example: R will tell us if the expression is TRUE or FALSE
6 < 5
```

```{r}
# s(7b) Write an expression with a logical operator that will return FALSE

```


################################
# Variable Assignment
################################

### A variable allows you to store a value (e.g. a number) or an object (e.g. a function description). 
### Later you can easily access the value or object stored within this variable
### Reminder: R is case sensitive

```{r}
# Example: Assign the value 25 to x

x <- 25

# Print out the value of the variable x
x
```

### You can view your variables in your Environment tab in RStudio

```{r}
# (s8) Assign the value of the square root of 144 to the variable my_var

# print out my_var

```

```{r}
# (s9) Assign the value of the 10 to the variable y

# print out y

```

###  You can use arthmetic operators and functions on variables if they are NUMBERIC variables

```{r}
# (s10) Subtract my_var from y and divide by 4 (Hint: your answer should be -0.5)

```

```{r}
# (s11) Assign the above equation to the variable my_ans and print it out

```

```{r}
# (s12) Assign the string "hello world" to the the variable char

```

```{r}
# (s13) Add char and y together (Hint:you will get an error, why?)

```

#############################
# Basic Data Types
############################

* Decimal values like 5.88 are called numerics
* Whole numbers like 5 are called integers. Integers are also numerics
* Boolean values (e.g., TRUE or FALSE) are called logical
* Text (or string) values are called characters (e.g. your variable 'char')

### We can check the data type of variable using the class() function

```{r}
# Example: Declare variables of different types

numeric <- 25
character <- "data science"
logical <- TRUE #note the color change of the text

```

```{r}
# (s14) Check the class of variables above using class() (Hint: class(numeric))

```

#############################
# Vectors
#############################

### Vectors are one-dimension arrays that can hold numeric, character, or logical data
### In R, you use the combine function, c(), with the elements separated with commas

```{r}
# Example: Creating vectors
numeric_vector <- c(41, 66, 23, 3)

# print numeric vector
numeric_vector

# check the class
class(numeric_vector)
```

```{r}
# (s15) Create, print, and check the class of a vector of your three favorite TV shows (Hint: don't forget the quotation marks) and assign it the variable tv

```

```{r}
# (s16) Create, print, and check the class of a vector named uva with the answers (True/False) to the following questions:
# UVA is located in Charlottesville.
# George Washington founded UVA.
# Data science is a useful field.


```

### We can add names to the data in our vectors using the names() function

```{r}
# Example: Create a vector
fruit <- c("apples", "red")

# Assign names to fruit vector
names(fruit) <- c("Fruit", "Color")

# Print to console
fruit
```

### Make this process more efficient by creating a vector of the names you want to use

```{r}
# (s17) Create a vector of numbers 1-7 (Hint: your can use the ':' operator, 1:7) and assign to the variable days_number

# Create a vector of the days of the week in order and save it to a variable called days_name

# Now assign the names of the day to the day number

```

### Arithmetic calcuations of vectors
* takes elements list-wise
* c(1, 2, 3) + c(4, 5, 6) = c(1 + 4, 2 + 5, 3 + 6)
* can do calculations with variables that represent vectors

```{r}
# Create 2 vectors and multiple them together

```

### Vector Selection
* use square brackets ([]) to select a specific element(s) out of your vector
* note: unlike other programming languages, the first element in a R vector is 1 (compared to 0)

```{r}
# Example
streets <- c('ivy', 'emmet', 'jpa', 'main')
streets[2]
```

```{r}
# (s18) Use your days_name vector and select the element "Thursday"

```

```{r}
# Example: Select multiple elements

# select the first 3 elements
streets[c(1, 2, 3)]
# =
streets[1:3]

# select the first and last elements (note the use of c())
streets[c(1,4)]

# select the first 2 elements and the last element
streets[c(1:2, 4)]
```

### Element selection can also be done using the names of elements instead of their numeric positions

```{r}
#Example:
# Vector of street numbers
street_numbers <- c('Street 1', 'Street 2', 'Street 3', 'Street 4')

# Assign street numbers as names to Streets
names(streets) <- street_numbers

# Select Street 1
streets['Street 1']

# Select Street 1 and Street 4
streets[c('Street 1', 'Street 2')]
```

# Vectors & logical operators 
### We can use logical operators to compare elements in a vector to a number (or == a string)
```{r}
#Example
c(105, 21, 67, 3, 88) > 45

# or we can use a vector saved to a variable

my_vector <- c(105, 41, 67, 3, 88)

my_vector > 45
```

```{r}
# s(1) Given the vector below, determine which elements are positive and save them to a vector called positive

c(-20, -0.5, 20, 0, 8, 3)

```

### If you place a logical vector inside square brackets, R will only return the elements that respond to TRUE

```{r}
# create a vector
v <- c(1, 2, 3, 4, 5, 6, 7)
names(v) <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday')

# logical vector
logical <- v > 3

# subset out the values that equal TRUE
thursday <- v[logical]
thursday
```